Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-15 Manuscript/manuscript.tex
Update text

\author{Josef Melcr} 
\email{melcr@uochb.cas.cz} 
%%\homepage[]{https://jmelcr.github.io/}
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\author{Pavel Jungwirth} 
%%\homepage[]{http://jungwirth.uochb.cas.cz/}
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101 
Tampere, Finland} 
%SAMULI: We need to put tiago because the experimental acyl chain order parameters
\author{Tiago Ferreira}
\affiliation{Halle} 
\author{O. H. Samuli Ollila} 
\email{samuli.ollila@helsinki.fi} 
%%\homepage[]{Your web page} 
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}

@@

The NMRlipids project \url{nmrlipids.blogspot.fi} has recently demonstrated
that such controversies can be resolved by using the headgroup order parameters
of phosphatidylcholine (PC) lipids \cite{catte16, nmrlipids_proj4}, which can
be related to cation binding affinity to lipid bilayers using the electrometer
concept \cite{akutsu81,altenbach84,seelig87}. The main advantage of the approach
is the direct comparison between experimental and calculated order parameters,
which reduced the ambiguity arising from the interpretation of the data.
Unfortunately, none of the readily available force fields was sufficiently accurate to correctly reproduce the
cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
mixtures with negatively charge PS lipids \cite{nmrlipids_proj4}.
In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
to zwitterionic 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC) lipid bilayer
by implicitly including electronic polarizability using the electronic continuum
correction (ECC) \cite{leontyev09}. The good agreement between the resulting ECC-POPC
model and experiments enable detailed interpretation of calcium binding details to POPC lipid bilayer.

@@

Here, we extend the ECC approach also to negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
The quality of the newly developed ECC-POPS force field parameters is
evaluated against previously published X-ray scattering form factors\cite{kucerka14}
and NMR \cite{roux90,nmrlipids_proj4} data for lipid headgroup C--H bond order parameters in various ionic conditions,
as well as against acyl chain C--H bond order parameters reported in this work.
as well as 
not only describes the interactions of PS with ions more accurately, 
but also improves the structure of the PS head group 
and its interactions with PC in mixed lipid bilayers. 
The results pave the way to more realistic MD simulations 
of both neutral and negatively charged lipid bilayers 
for a wide range of applications.

%Here, we demonstrate that the same approach 
%significantly improves the description of calcium binding 
%also to the negatively charged PC:PS (5:1) bilayers. 
%In correspondence with the Ref.~\citenum{nmrlipids_proj4},
%we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
%in PC and PS lipids 
%to quantify the binding affinity of the calcium cations
%and the induced structural changes
%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
%To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
%and to mark the improvement after treating electronic polarizability, 
%we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}.

@@

In the reference simulations with the standard Amber force field,
the Lipid14 parameters~\cite{dickson14} for POPC, the Lipid17 parameters for POPS \cite{lipid17-future},
the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} were used.
The previously generated \cite{botan15} Lipid14 parameters in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
The Lipid17 parameters were obtained from AmberTools18 \cite{amber18} 
and converted to Gromacs format using acpype \cite{acpype}.  
The ion model by Dang and coworkers~\cite{smith94,chang1999,dang2006} was used with Amber lipids because
the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{nmrlipids_proj4}.
In the ECC-lipid simulations,
the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
the ECC-POPS parameteres derived in this work (section \ref{section:ecc}), 
the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} were used.
The SPC/E water model was selected because its lower dielectric constant is consistent with the
ECC concept \cite{leontyev11,leontyev14}.
\begin{table}[tbp]
  \caption{Simulation parameters}
  \label{tbl:mdpar}
  \begin{tabular}{ll}
    simulation property & parameter   \\
    \hline
    time-step           & 2~fs         \\
    equilibration time  & 50~ns  \\
    total simulation time     & 1~$\mu$s  \\
    temperature         & 298~K       \\
    thermostat          & v-rescale  \cite{bussi07}   \\
    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
    long-range electrostatics & PME  \cite{darden93}  \\
    cut-off scheme      & Verlet \cite{Pall13}      \\
    Coulomb and VdW cut-off & 1.0~nm \\
    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
    constraints for water & SETTLE  \cite{miyamoto92} \\
    \hline
  \end{tabular}
\end{table}

The C--H bond order parameters were calculated directly from Eq. \ref{OPequation}.
Average over the time for individual lipids was calculated first and
the standard error of the mean of the average over lipids was then used as the
error estimate \cite{botan15,ollila16,nmrlipids_proj4}.
Python program ({\tt scripts/calcOrderParameters.py}) that uses the
MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
The ion number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.
The salt concentrations in buffer before solvating the lipids, reported in the used experimental data set \cite{roux90},
were calculated as $[{\rm salt}]=N_{\rm c} \times[{\rm water}]\,/\,N_{\rm w}$,
where $N_{\rm c}$ is the number of cations in simulation, $[{\rm water}]$\,=\,55.5~M and $N_{\rm w}$
is the number of water molecules in the simulation.
As discussed in our previous work \cite{nmrlipids_proj4}, the hydration levels of
multilamellae are expected to be sufficiently similar in the used simulations and reference experiments \cite{roux90}.

@@

The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,nmrlipids_proj4}.
Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
order parameters also from the acyl chain region using the R-PDLF experiments as described previously \cite{ferreira13}.
We first determined the order parameters of each peak corresponding different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}
\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
    compared with experiments at 298~K. \citep{nmrlipids_proj4}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
  \todo{We should use the same color coding for the models as in ECC-POPC paper, now it is opposite.} \\
  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
\end{figure}

To evaluate the effect of ECC on the structural properties of POPS lipid bilayer,
we compared the X-ray scattering form factors, C--H bond order parameters and area per molecules
between simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??},
ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}).
The larger acyl chain order parameters, deviation in the form factor and smaller area per molecule 
in the Lipid17 simulation with Dang ions when compared with experiments indicates that this simulation
predicts too dense lipid bilayer. In previous study \cite{nmrlipids_proj4}, Lipid17 simulated with {\AA}qvist ions~\cite{aqvist90}
gave larger area per molecule (57 \AA$^2$), probably due to the weaker counterion binding affinity.
The Dang ions are used to here to avoid the artificial pairing of ions and condensation at larger salt
concentrations observed with {\AA}qvist \cite{kohagen16, chen07, nmrlipids_proj4} when analysing the effect of ECC. 
The scattering form factors and acyl chain order parameters from ECC-POPS are in good agreement with
experiments (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
are within each others error bars. 
%In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
%we cannot decide, which of the values is more reliable.

@@

Despite the success in simulating bilayer dimensions and acyl chain conformations,
lipid force fields have typically problems in capturing the correct glycerol backbone and
headgroup order parameters and conformations \cite{botan15,ollila16,nmrlipids_proj4}.
This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
and {\AA}qvist \cite{nmrlipids_proj4} ions. The values from ECC-POPS simulation are closer,
but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
Especially, the experimentally observed more negative values of the $\beta$-carbon
and the larger forking (difference of order parameters between hydrogens attached to the
same carbon) of the $\alpha$-carbon in POPS than in POPC \cite{nmrlipids_proj4}
are not reproduced by neither of the models here. These differences are previously related
to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
dihedral, which is \todo{the dihedral should be plotted and sentence finished.}

@@

In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
acyl chain conformations in good agreement with experiments, but has room for
improvement in the glycerol backbone and headgroup region. Therefore,
its structural accuracy is in the same level as in other state-of-the-art
lipid force fields \cite{botan15, ollila16, Pluhackova2016, nmrlipids_proj4}.

@@

Binding affinity of monovalent ions to lipid bilayers depends strongly on
force field parameters \cite{catte16,melcr18,nmrlipids_proj4},
which is the case also here as the sodium binding affinity to
POPS lipid bilayer is stronger in Lipid17 simulation with Dang ions than in
ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
The cation binding affinities to lipid bilayers in simulations have been previously evaluated
against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,nmrlipids_proj4},
which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
it is less well undestood and the addition of cations may cause phase transitions in negatively charged
bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
mixtures\cite{roux90,nmrlipids_proj4}.
Following our previous work \cite{nmrlipids_proj4}, we evaluate the counterion binding affinity
to lipid bilayer in simulations containing PS using two different approaches.

@@

%Unlinke in pure PC bilayers,
%Interpreting the changes of the head group order parameters of 
%mixed PC:PS lipid bilayers 
%from both simulations and experiments
%is complicated by the presence of counterions. \cite{nmrlipids_proj4}

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.
  SAMULI: I think that these are not promised in the current text and addition is not necessary.} \\
\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?
  SAMULI: I feel now that we should keep the figure as it is. Also the POPC:POPS (5:1) results are now referred in the text.
  However, if something is moved to SI, it should be POPS in main text, mixture in SI.}
\end{figure}

@@

%is the
%is not generally agreed on between simulations 
%\citep{bockmann03,sachs04,berkowitz06,cordomi09}
%and experiments 
%\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
%We attempted to find a model, 
%which would successfully interpret the experiments on binding of \ce{Na^+}
%to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
%in our works \citep{catte16, nmrlipids_proj4}.
%After testing the currently available force fields for phospholipids,
%we concluded
%that the interactions are generally overestimated in magnitude in almost all models 
%but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
%This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
%when used with the model of ions by \citet{aqvist90}. 
%However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
%the model overestimates the binding affinity of \ce{Na^+}
%measured with lipid electrometer concept. \citep{melcr18}
%In total, these results suggest that improvements 
%in the lipid parameters are required for more accurate interactions even with monovalent cations. 
%\citep{catte16, melcr18, nmrlipids_proj4}

>>>>

\author{Josef Melcr} 
\email{melcr@uochb.cas.cz} 
%%\homepage[]{https://jmelcr.github.io/}
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\author{Tiago Ferreira}
\affiliation{NMR group - Institut for Physics, Martin-Luther University Halle-Wittenberg} 
\author{Pavel Jungwirth} 
%%\homepage[]{http://jungwirth.uochb.cas.cz/}
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101 
Tampere, Finland} 
%SAMULI: We need to put tiago because the experimental acyl chain order parameters
\author{O. H. Samuli Ollila} 
\email{samuli.ollila@helsinki.fi} 
%%\homepage[]{Your web page} 
\affiliation{Institute of Organic Chemistry and Biochemistry, 
Academy of Sciences of the Czech Republic,  
Prague 6, Czech Republic} 
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}

@@

The NMRlipids project \url{nmrlipids.blogspot.fi} has recently demonstrated
that such controversies can be resolved by using the headgroup order parameters
of phosphatidylcholine (PC) lipids \cite{catte16, NMRlipidsIV}, which can
be related to cation binding affinity to lipid bilayers using the electrometer
concept \cite{akutsu81,altenbach84,seelig87}. The main advantage of the approach
is the direct comparison between experimental and calculated order parameters,
which reduced the ambiguity arising from the interpretation of the data.
Unfortunately, none of the readily available force fields was sufficiently accurate to correctly reproduce the
cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
mixtures with negatively charge PS lipids \cite{NMRlipidsIV}.
In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
to zwitterionic 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC) lipid bilayer
by implicitly including electronic polarizability using the electronic continuum
correction (ECC) \cite{leontyev09}. The good agreement between the resulting ECC-POPC
model and experiments enable detailed interpretation of calcium binding details to POPC lipid bilayer.

@@

Here, we extend the ECC approach also to negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
We also complement the available experimental data for the the force field quality
evaluation by measuring the acyl chain C--H bond order parameters of POPS in bilayer using natural abundance $^{13}$C NMR.
In addition to the acyl chain order parameters measured here, the quality of the newly developed ECC-POPS force field parameters is
evaluated using previously published C--H bond order parameters of glycerol backbone and headgroup in various ionic
conditions and lipid molar fractions \cite{roux90,NMRlipidsIV}, as well as X-ray scattering form factors\cite{kucerka14}.
The overall improvement of the force field accuracy upon applying ECC
%not only describes the interactions of PS with ions more accurately, 
%but also improves the structure of the PS head group 
%and its interactions with PC in mixed lipid bilayers. 
%The results
paves the way to more realistic MD simulations 
of both neutral and negatively charged lipid bilayers 
for a wide range of applications.

%Here, we demonstrate that the same approach 
%significantly improves the description of calcium binding 
%also to the negatively charged PC:PS (5:1) bilayers. 
%In correspondence with the Ref.~\citenum{NMRlipidsIV},
%we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
%in PC and PS lipids 
%to quantify the binding affinity of the calcium cations
%and the induced structural changes
%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
%To directly connect to the previous works \citep{catte16, NMRlipidsIV}
%and to mark the improvement after treating electronic polarizability, 
%we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}.

@@

In the reference simulations with the standard Amber force field,
the Lipid14 parameters~\cite{dickson14} for POPC, the Lipid17 parameters for POPS \cite{lipid17-future},
the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} were used.
The previously generated \cite{botan15} Lipid14 parameters in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
The Lipid17 parameters were obtained from AmberTools18 \cite{amber18} 
and converted to Gromacs format using acpype \cite{acpype}.  
The ion model by Dang and coworkers~\cite{smith94,chang1999,dang2006} was used with Amber lipids because
the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{NMRlipidsIV}.
In the ECC-lipid simulations,
the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
the ECC-POPS parameteres derived in this work (section \ref{section:ecc}), 
the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} were used.
The SPC/E water model was selected because its lower dielectric constant is consistent with the
ECC concept \cite{leontyev11,leontyev14}.
\begin{table}[tbp]
  \caption{Simulation parameters}
  \label{tbl:mdpar}
  \begin{tabular}{ll}
    simulation property & parameter   \\
    \hline
    time-step           & 2~fs         \\
    equilibration time  & 50~ns  \\
    total simulation time     & 1~$\mu$s  \\
    temperature         & 298~K       \\
    thermostat          & v-rescale  \cite{bussi07}   \\
    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
    long-range electrostatics & PME  \cite{darden93}  \\
    cut-off scheme      & Verlet \cite{Pall13}      \\
    Coulomb and VdW cut-off & 1.0~nm \\
    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
    constraints for water & SETTLE  \cite{miyamoto92} \\
    \hline
  \end{tabular}
\end{table}

The C--H bond order parameters were calculated directly from Eq. \ref{OPequation}.
Average over the time for individual lipids was calculated first and
the standard error of the mean of the average over lipids was then used as the
error estimate \cite{botan15,ollila16,NMRlipidsIV}.
Python program ({\tt scripts/calcOrderParameters.py}) that uses the
MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
The ion number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.
The salt concentrations in buffer before solvating the lipids, reported in the used experimental data set \cite{roux90},
were calculated as $[{\rm salt}]=N_{\rm c} \times[{\rm water}]\,/\,N_{\rm w}$,
where $N_{\rm c}$ is the number of cations in simulation, $[{\rm water}]$\,=\,55.5~M and $N_{\rm w}$
is the number of water molecules in the simulation.
As discussed in our previous work \cite{NMRlipidsIV}, the hydration levels of
multilamellae are expected to be sufficiently similar in the used simulations and reference experiments \cite{roux90}.

@@

The X-ray scattering form factors and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
We first determined the order parameters of each peak corresponding different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}
\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
    compared with experiments at 298~K. \citep{NMRlipidsIV}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
  \todo{We should use the same color coding for the models as in ECC-POPC paper, now it is opposite.} \\
  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
\end{figure}

The Lipid17/Dang simulation give larger acyl chain order parameters and smaller area per molecule than the
experimental values (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}),
indicating that this simulation predicts too dense POPS lipid bilayer. This is also in line with
the discrepancy observed in the X-ray form factor between this simulation and experiment (Fig. \ref{simVSexpNOions_POPS}).
In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
gave a larger area per molecule (57 \AA$^2$) for POPS being closer to the experimental value,
which is probably due to the weaker counterion binding affinity of {\AA}qvist ions.
However,  {\AA}qvist ions exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of POPS lipid bilayer. 
The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
are within each others error bars.

% In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
%we cannot decide, which of the values is more reliable.

@@

Despite the success in simulating bilayer dimensions and acyl chain conformations,
lipid force fields have typically problems in capturing the correct glycerol backbone and
headgroup order parameters and conformations \cite{botan15,ollila16,NMRlipidsIV}.
This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
and {\AA}qvist \cite{NMRlipidsIV} ions. The values from ECC-POPS simulation are closer,
but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
Especially, the experimentally observed more negative values of the $\beta$-carbon
and the larger forking (difference of order parameters between hydrogens attached to the
same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
are not reproduced by neither of the models here. These differences are previously related
to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
dihedral, which is \todo{the dihedral should be plotted and sentence finished.}

@@

In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
acyl chain conformations in good agreement with experiments, but has room for
improvement in the glycerol backbone and headgroup region. Therefore,
its structural accuracy is in the same level as in other state-of-the-art
lipid force fields \cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
    in ECC-lipid (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.
  SAMULI: I think that these are not promised in the current text and addition is not necessary.} \\
\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?
  SAMULI: I feel now that we should keep the figure as it is. Also the POPC:POPS (5:1) results are now referred in the text.
  However, if something is moved to SI, it should be POPS in main text, mixture in SI.}
\end{figure} 
Binding affinity of monovalent ions to lipid bilayers depends strongly on
force field parameters \cite{catte16,NMRlipidsIV},
which is the case also here as the sodium counterion binding affinity to
POPS lipid bilayer is stronger in Lipid17/ Dang simulation than in
ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
The cation binding affinities to lipid bilayers in simulations have been previously evaluated
against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
it is less well undestood and the addition of cations may cause phase transitions in negatively charged
bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
mixtures\cite{roux90,NMRlipidsIV}.
Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
to lipid bilayer in simulations containing PS using two different approaches.

@@

%Unlinke in pure PC bilayers,
%Interpreting the changes of the head group order parameters of 
%mixed PC:PS lipid bilayers 
%from both simulations and experiments
%is complicated by the presence of counterions. \cite{NMRlipidsIV}

@@

%is the
%is not generally agreed on between simulations 
%\citep{bockmann03,sachs04,berkowitz06,cordomi09}
%and experiments 
%\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
%We attempted to find a model, 
%which would successfully interpret the experiments on binding of \ce{Na^+}
%to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
%in our works \citep{catte16, NMRlipidsIV}.
%After testing the currently available force fields for phospholipids,
%we concluded
%that the interactions are generally overestimated in magnitude in almost all models 
%but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
%This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
%when used with the model of ions by \citet{aqvist90}. 
%However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
%the model overestimates the binding affinity of \ce{Na^+}
%measured with lipid electrometer concept. \citep{melcr18}
%In total, these results suggest that improvements 
%in the lipid parameters are required for more accurate interactions even with monovalent cations. 
%\citep{catte16, melcr18, NMRlipidsIV}
